\documentclass{article}

\usepackage[
	eientei,
	+maths,
	+tikz,
	categoryStyle = script,
	scriptStyle = rsfs
]{kappak}
\usepackage{paralist}
\usepackage{stmaryrd}

\title{Homotopy Theory of Strict $\omega$-Categories}
\author{Unofficial notes from the lectures of François Métayer \\ \LaTeX ed by Cédric HT}
\date{25 -- 28 September 2017}

\newcategory{\Poly}{}{P}{oly}

\newcommand{\Weq}{\frakWW}
\newcommand{\Fib}{\mathfrak{Fib}}
\newcommand{\Cof}{\mathfrak{Cof}}

\begin{document}

\maketitle

\begin{definition*}
This document is unofficial notes taken from the lectures ``Homotopy Theory of Strict $\omega$-Categories'' given by François Métayer during the conference ``Categories in Homotopy Theory and Rewriting'', held at CIRM in September 2017. It can contain many typos and mistakes, and some notations have been changed from the original lecture.
\end{definition*}

\section{Model categories}

Throughout this section, we assume the ambient category $\catCC$ to be complete and cocomplete. We make use of the grammatical composition convention, meaning that the composite of $A \xrightarrow{f} B \xrightarrow{g} C$ is written $fg$.

\paragraph*{Lifting properties. } Consider the following diagram:
\[ \begin{tikzpicture}[inline/.style = {fill = white, inner sep = 2.5pt}]
        \matrix (m) [matrix of math nodes, row sep = 2em, column sep = 3em, text height = 1.5ex, text depth = 0.25ex] {
            \bullet & \bullet \\
            \bullet & \bullet \\
        };
        \path[->]
            (m-1-1) edge                              (m-1-2)
            (m-1-1) edge          node [left]   {$f$} (m-2-1)
            (m-1-2) edge          node [right]  {$g$} (m-2-2)
            (m-2-1) edge                              (m-2-2)
            (m-2-1) edge [dotted] node [inline] {$k$} (m-1-2);
    \end{tikzpicture} \]
We say that $f$ has the left lifting property with respect to $g$ if such a morphism $k$ making the two triangles commute exists. We also say that $g$ has the right lifting property with respect to $f$, and denote it by $f \boxslash g$. If $\frakII$ is a class of morphism, let ${}^\boxslash \frakII = \{ f \in \hom \catCC \mid f \boxslash i, \forall i \in \frakII \}$ and similarily for $\frakII^\boxslash$. The class ${}^\boxslash \frakII$ of $\frakII$-injective morphism is closed under pushouts, sums, retracts, and transfinite composition.

\paragraph*{Small objects. } We say that an object $A$ is small if the representable $\catCC (A, -)$ preserves countable directed colimits. If $\frakII$ is a class of morphism whose domain are all small, we can apply the following:

\begin{theorem}[Small object argument]
Under the conditions above, any morphism $f \in \frakII$ can be factorized as $f = i p$, with $p \in {}^\boxslash \frakII$ and $i \in {}^\boxslash (\frakII^\boxslash)$.
\end{theorem}

\paragraph*{Weak factorization systems. } A weak factorization system is a pair $(\frakLL, \frakRR)$ of classes of morphisms, such that: 
\begin{inparaenum}[1.]
  \item each morphism $f$ of $\catCC$ factors as $f = rl$ with $r \in \frakRR$ and $l \in \frakLL$;
  \item $\frakLL = {}^\boxslash \frakRR$;
  \item $\frakRR = \frakLL^\boxslash$.	
\end{inparaenum}

\begin{proposition}
If $\frakII$ has small domains, then $\left( {}^\boxslash (\frakII^\boxslash), \frakII^\boxslash \right)$ is a weak factoriz	ation system.
\end{proposition}

\paragraph*{Model structure. } A model structure on $\catCC$ is the data of three classes of morphisms: $\Weq$ (weak equivalences), $\Cof$ (cofibrations), and $\Fib$ (fibrations), such that:
\begin{enumerate}
	\item $\Weq$ has the ``3 for 2'' property (also known as ``2 out of 3''): among $f$, $g$, $fg$, if two are in $\Weq$ then so is the third;
	\item $\Weq$, $\Cof$, and $\Fib$ are closed under retracts;
	\item $\Weq \cap \Cof \subseteq {}^\boxslash \Fib$, and $\Weq \cap \Fib \subseteq \Cof^\boxslash$;
	\item we have two weak factorization systems: $(\Weq \cap \Cof, \Fib)$, and $(\Cof, \Weq \cap \Fib)$.
\end{enumerate}

This is the classical definition, that we shall not use. The following one is proved equivalent in \cite{MR1780498}: a model structure on $\catCC$ is the data of two classes of morphisms: $\Weq$ (weak equivalences) and $\frakII$ (generating cofibrations), such that:
\begin{enumerate}
	\item $\Weq$ has the ``3 for 2'' property and is stable under retracts;
	\item $\frakII^\boxslash \subseteq \Weq$;
	\item $\Weq \cap {}^\boxslash (\frakII^\boxslash)$ is stable under pushouts and transfinite composition;
	\item the ``solution set condition'' holds: for $i \in \frakII$, let $J_i$ be the class of morphisms $j$ that factor an outer square as follows:
		\[ \begin{tikzpicture}[inline/.style = {fill = white, inner sep = 2.5pt}]
	        \matrix (m) [matrix of math nodes, row sep = 2em, column sep = 3em, text height = 1.5ex, text depth = 0.25ex] {
	            \bullet &         & \bullet \\
	                    & \bullet &         \\
	                    & \bullet &         \\
				\bullet &         & \bullet \\
	        };
	        \path[->]
	            (m-1-1) edge                                       (m-1-3)	            	            
	            (m-1-1) edge          node [left]   {$i$}          (m-4-1)
	            (m-1-3) edge          node [right]  {$w \in \Weq$} (m-4-3)
	            (m-4-1) edge                                       (m-4-3)
	            (m-1-1) edge [dotted]                              (m-2-2)
	            (m-2-2) edge [dotted]                              (m-1-3)
	            (m-2-2) edge [dotted] node [left]   {$j$}          (m-3-2)
	            (m-4-1) edge [dotted]                              (m-3-2)
	            (m-3-2) edge [dotted]                              (m-4-3);
	    \end{tikzpicture} \]
	    we then ask that $J_i \subseteq \Weq \cap {}^\boxslash (\frakII^\boxslash)$.
\end{enumerate}

\begin{theorem}[Jeff Smith]
Under the above conditions, letting $\Cof = {}^\boxslash (\frakII^\boxslash)$ and $\Fib = (\Weq \cap \Cof)^\boxslash$ yields a model structure in the classical definition.
\end{theorem}


\section{$\omega$-categories}

\paragraph*{Globular sets. } Let $\bbGG$ be the category of globes, having objects the natural numbers, and morphisms $\sigma, \tau : n \longrightarrow n+1$, for all $n \in \bbNN$, subject to the following globularity conditions:
\[ \left\{ \begin{array}{l}
	\sigma \sigma = \sigma \tau \\ \tau \sigma = \tau \tau .
\end{array} \right. \]
The category of globular sets is defined as $\hat{\bbGG} = [ \bbGG^\op, \Set ]$.

\paragraph*{$\omega$-categories. } An $\omega$-category $\catCC$ is a globular set
\[ \catCC_0 \xleftarrow{s, t} \catCC_1 \xleftarrow{s, t} \cdots \xleftarrow{s, t} \catCC_n \xleftarrow{s, t} \catCC_{n+1} \xleftarrow{s, t} \cdots , \]
with:
\begin{enumerate}
	\item compositions: for $x, y \in \catCC_n$ with $tx = sy$, we have a composition $x \circ_{n-1} y \in \catCC_n$;
	\item identity cells: for $x \in \catCC_n$ we have $\id (x) \in \catCC_{n+1}$ being left and right neutral for $\circ_n$;
\end{enumerate}
such that:
\begin{enumerate}
	\item composition is associative;
	\item the exchange rule holds:
		\[ (x \circ_i y) \circ_j (z \circ_i w) = (x \circ_j z) \circ_i (y \circ_j w) \]
		whenever those expressions are defined.
\end{enumerate}
Let $\Cat_\omega$ be the category of strict $\omega$-categories and natural transformations that preserve composition and identities, also called $\omega$-functors.

For $\catCC \in \Cat_\omega$ and $x, y \in \catCC_0$, the $\hom$ $[x, y]$ is also an $\omega$-category. For $u \in \catCC_{n+1}$ such that $s^0 u = x$ (iterated source at dimension $0$) and $t^0 = y$, we have a cell $[u] \in [x, y]_n$.

The forgetful functor $\Cat_\omega \longrightarrow \hat{\bbGG}$ has a left adjoint. Let $O[n]$ be the free $\omega$-category over the representable $\bbGG (-, n)$. Let $\partial O [n]$ be the boundary of $O [n]$, i.e. the free $\omega$-category over $\bbGG (-, n)$ with the $n$-cell removed. We have canonical inclusions $i_n : \partial O [n] \longincl O [n]$, and let $\frakII = \{ i_n \mid i \in \bbNN \}$.

\begin{definition}
This definition is made by mutual coinduction.
\begin{itemize}
	\item For $n \geq 1$, a $n$-cell $u : x \longrightarrow y$ is reversible if there is a cell $\barU : y \longrightarrow x$ such that $u \circ_{n-1} \barU \sim \id (x)$ and $\barU \circ_{n-1} u \sim \id (y)$;
	\item For $n \geq 1$, two parallel $n$-cells $f$ and $g$ are $\omega$-equivalent, denoted by $f \sim g$, if there exists a reversible $(n+1)$-cell $u : f \longrightarrow g$.
\end{itemize}
\end{definition}

For instance, identities are reversible cells. An $\omega$-functor $F : \catCC \longrightarrow \catDD$ is a weak equivalence if
\begin{enumerate}
	\item $F_0 : \catCC_0 \longrightarrow \catDD_0$ is surjective up to $\sim$;
	\item for parallel $x, y \in \catCC_n$, of there is $w : Fx \longrightarrow Fy$, then there exists $v : x \longrightarrow y$ such that $Fv \sim w$.
\end{enumerate}
Let $\Weq$ be the class of such weak equivalences.

\begin{theorem}
The category $\Cat_\omega$ together with $\Weq$ as the class of weak equivalences and $\frakII$ as the class of generating cofibrations is a model category.
\end{theorem}


\section{Cofibrant objects and polygraphs}

\paragraph*{Free category over a cellular extension. } Let $\bbGG|_n$ be the category $\bbGG$ truncated at dimension $n$, i.e. the full subcategory generated by objects $0, \ldots, n$. We call elements of $\widehat{\bbGG|_n}$ $n$-globular sets. Consider the following pullback:
\[ \pullbackdiagram{\Cat_n^+}{\widehat{\bbGG|_{n+1}}}{\Cat_n}{\widehat{\bbGG|_n} ,}{}{}{}{} \]
where the cospan contains the forgetful functors. An object of $\Cat_n^+$ is then a $n$-category $\catCC$ endowed with a cellular extension $S_{n+1} \xrightarrow{s, t} \catCC_n$ satisfying the globularity condition. The outer square
\[ \begin{tikzpicture}[inline/.style = {fill = white, inner sep = 2.5pt}]
        \matrix (m) [matrix of math nodes, row sep = 2em, column sep = 3em, text height = 1.5ex, text depth = 0.25ex] {
           	\Cat_{n+1} &          &                        \\
           	           & \Cat_n^+ & \widehat{\bbGG|_{n+1}} \\
			           & \Cat_n   & \widehat{\bbGG|_n} ,   \\
        };
        \path[->]
            (m-1-1) edge [out =   0, in = 135]                               (m-2-3)
            (m-1-1) edge [out = -90, in = 160]                               (m-3-2)
            (m-1-1) edge [dotted]              node [inline] {$\exists ! W$} (m-2-2)
            (m-2-2) edge                                                     (m-2-3)
            (m-2-2) edge                                                     (m-3-2)
            (m-2-3) edge                                                     (m-3-3)
            (m-3-2) edge                                                     (m-3-3);
		\path
			(m-2-2) edge [-, draw = white]     node [inline] {\mbox{\LARGE{$\lrcorner$}}} (m-3-3);
    \end{tikzpicture} \]
factors uniquely trough a forgetful functor $W : \Cat_{n+1} \longrightarrow \Cat_n^+$ that forgets the composition $\circ_n$. It has a left adjoint $L$, so that from a cellular extension $\catCC^+ = (S_{n+1} \xrightarrow{s, t} \catCC)$ in $\Cat_n^+$ we can build a free $L \catCC^+ \in \Cat_{n+1}$.

\paragraph*{Polygraph. } A $0$-polygraph is a set endowed with its identity:
\[ \begin{tikzpicture}
    \matrix (m) [matrix of math nodes,
            row sep = 3 em,
            column sep = 4 em,
            text height = 1.5 ex,
            text depth = 0.25 ex] {
        P_0 \\
        P_0^* \\
    };
    \path[->]
    	(m-1-1) edge node [left] {$\id$} (m-2-1);
\end{tikzpicture} \]
By induction, a $(n+1)$-polygraph is a $n$ polygraph with a cellular extension $P_{n+1} \xrightarrow{s, t} P_n^*$, which generate a free $(n+1)$-category, whose set of $(n+1)$-cells is denoted by $P_{n+1}^*$:
\[ \begin{tikzpicture}
    \matrix (m) [matrix of math nodes,
            row sep = 3 em,
            column sep = 4 em,
            text height = 1.5 ex,
            text depth = 0.25 ex] {
        P_0   & P_1   & \cdots & P_n   & P_{n+1}   \\
        P_0^* & P_1^* & \cdots & P_n^* & P_{n+1}^* \\
    };
    \path[->]
    	(m-1-1) edge                       (m-2-1)
    	(m-1-2) edge                       (m-2-2)
    	(m-1-4) edge                       (m-2-4)
    	(m-1-5) edge                       (m-2-5)
    	
    	(m-1-2) edge node [above] {$s, t$} (m-2-1)
    	(m-1-3) edge node [above] {$s, t$} (m-2-2)
    	(m-1-4) edge node [above] {$s, t$} (m-2-3)
    	(m-1-5) edge node [above] {$s, t$} (m-2-4)
    	
    	(m-2-2) edge node [below] {$s, t$} (m-2-1)
    	(m-2-3) edge node [below] {$s, t$} (m-2-2)
    	(m-2-4) edge node [below] {$s, t$} (m-2-3)
    	(m-2-5) edge node [below] {$s, t$} (m-2-4);
\end{tikzpicture} \]
A polygraph $P = (P_0, P_1, \ldots)$ has an underlying $\omega$-category, denoted by $P^*$. Morphisms of polygraphs are $\omega$-functors that map generators to generators. We denote by $\Poly_\omega$ the category of polygraphs and such morphisms.

\begin{proposition}
If $\catCC = P^*$ is a $\omega$-category generated by a polygraph, then $\catCC$ is cofibrant.
\end{proposition}
\begin{proof}[Sketch]
The successive truncations $\catCC|_n$ of $\catCC$ can be obtain by the following pushout:
\diagramarrows{}{>->}{>->}{}
\[ \pushoutdiagram{\coprod_{P_{n+1}} \partial O [n+1]}{\catCC|_n}{\coprod_{P_{n+1}} O [n+1]}{\catCC|_{n+1}}{}{}{\eta_n}{} \]
and since the left arrow is a cofibration, so is the right one. Then the initial inclusion map $\emptyset \longrightarrow \catCC$ is obtained via the following transfinite composite of cofibrations, and hence is a cofibration:
\[ \emptyset \longmono \catCC|_0 \stackrel{\eta}{\longmono} \catCC|_1 \longmono \cdots \longrightarrow \catCC \cong \colim_n \catCC|_n . \]
\end{proof}

Let $P \catCC$ be the normal resolution of an $\omega$-category $\catCC$, i.e. the natural polygraph such that $P \catCC_n = \catCC_n$.

\begin{proposition}
The normal resolution gives rise to an adjunction $(-)^* : \Poly_\omega \stackrel{\dashv}{\longleftrightarrow} \Cat_\omega : P$. Moreover, the counit $\epsilon_\catCC : (P \catCC)^* \longrightarrow \catCC$ is a trivial fibration, making the comonad $(P -)^*$ a cofibrant replacement.
\end{proposition}

\nocite{*}
\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}